Towards a Neural Co-Processor Which Restores Movement After Stroke: Modeling a Proof-of-Concept

### 2021-11-07 paper/article.tex
Paper progress

\subsection{Architecture Overview}
First, we present the architecture of our co-processor design. This design aims to solve two
fundamental challenges in using neural stimulation to improve external task performance.
First, neural networks exhibit long-running and highly nonlinear dependencies, necessitating
the need for a stimulation agent to account for far-distant effects of the stimulation it
applies. Second, with neurologically complex tasks, such as grasping, the mapping between
neural activity as-measured and the external task cannot be determined \textit{a priori}.
As a result, the co-processor must somehow learn what stimulation is appropriate for
aiding the user in the external task they are attempting to perform.

@@

There are some key differences between that approach and the EN architecture we
present here. Our EN's purpose is simply to train the CPN, so we pick an
architecture we find to be best suited to that job. First, our EN architecture
is not linear, through the use of a $tanh$ activation function. That allows us
to escape some of the limits in representational power of linear models, though
we didn't deeply test a strictly linear approach.
Second, LSTM cells are designed to better capture long-running network
dynamics, and in our case we found that they were crucial for that reason.
Compare that to the simpler cells commonly used in vanilla RNNs, which
yield functionally the same approach as the linear state space model, less
the use of a nonlinear activation function.

\subsection{Simulation Overview}
We demonstrate our approach here using a simulated grasping circuit. A detailed
simulation such as this allows us to explore some of the critical details in
how to train a co-processor of this design. By first doing such exploration in
simulation, we are able to rapidly and cheaply iterate on our design, prior to
any \textit{in vivo} experiments. In order for the simulation to be
admissible, we need to ensure it has some properties that allow it to strongly
indicate if our design is improving in a direction that will later allow for
real deployments. Otherwise, our design may be adapting to the pecularities of
the simulation, witheout becoming more useful.

@@

The stimulation model in our case adapts itself to the task of training our closed-
loop AI stimulation agent. For a neurologically complex task such as grasping, we
don't know \textit{a priori} what stimulation regimes to to explore while searching
for an optimal solution. Our stimulation model plays the role of generating learning
signals, via error backpropagation (<TODO? Citation?>), for our stimulation agent.
It continuously updates itself based on observations of brain activity, stimulation,
and external task queues, such as hand position.

@@

* Overview of test bed
** Michaels model \cite{michaels.mrnn} \cite{susillo.mrnn}.
** Lesion designs
*** Lesion examples, i.e. hand velocity more affected than shoulder velocity.
*** Disconnect modules vs lesion M1, and the effects
** Stimulation design
*** Spatio-temporal smoothing
** Observation model

>>>>

\subsection{Architecture Overview}
First, we present the architecture of our co-processor design. This design aims to solve two
fundamental challenges in using neural stimulation to improve external task performance.
First, neural networks exhibit long-running and nonlinear dependencies, necessitating
the need for a stimulation agent to account for far-distant effects of the stimulation it
applies. Second, with neurologically complex tasks, such as grasping, the mapping between
neural activity as-measured and the external task cannot be determined \textit{a priori}.
As a result, the co-processor must somehow learn what stimulation is appropriate for
aiding the user in the external task they are attempting to perform.

@@

There are some key differences between that approach and the EN architecture we
present here. Our EN's purpose is simply to train the CPN, so we pick an
architecture we find to be best suited to that job. First, our EN architecture
is not linear, through the use of a $tanh$ activation function. That allows us
to escape some of the limits in representational power of linear models, though
we didn't deeply test a strictly linear approach.
Second, LSTM cells are designed to better capture long-running network
dynamics, and in our case we found that they were crucial for that reason:
when using vanilla RNN cells, our co-processor wasn't able to train well.
Compare that to the simpler cells commonly used in vanilla RNNs, which
yield functionally the same approach as the linear state space model, less
the use of a nonlinear activation function.

\subsection{Simulation Overview}
We demonstrate our approach here using a simulated grasping circuit.
A detailed simulation such as this allows us to explore some of the critical details in
training a co-processor of this design. By first doing such exploration in
simulation, we are able to rapidly and cheaply iterate on our design, prior to
any \textit{in vivo} experiments. In order for the simulation to be
admissible, we need to ensure it has some properties that allow it to strongly
indicate if our design is improving in a direction that will later allow for
real deployments. Otherwise, our design may be adapting to the pecularities of
the simulation, without becoming more useful for the real world.

The simulated circuit, from Michaels et al. \cite{michaels.mrnn, susillo.mrnn}, was trained
to resemble the grasping circuits of monkey subjects engaged in a reach-to-grasp task. It
consists of a ``modular'' vanilla RNN, and a linear readout layer. The inputs are visual
features intended to capture the view the monkeys had during the task, i.e. 3D renderings of
the same objects which the monkeys grasped. The outputs are muscle length velocities for the
should, arm, and hand of the monkey during the trial. The natural velocities were captured
with a motion capturing glove, and the artificial neural network is trained to recapitulate
those grasping motions. See Fig. \ref{fig:michaels}.

The design of the circuit is intended to resemble a vision-to-grasp pipeline, essentially
representing the visual processing needed to reach the hand to the appropriate position for
the grasp, and to form the hand properly for grasping the particular shape of object. The
emergent dynamics of the artificial network's ``modules'', once train, correspond roughly
to the AIP, F5, and M1 portions of the monkey subjects' brains. Notably, the activity of
the module receiving the visual inputs resembles the AIP activity of the monkey subject
from the same trials. Likewise, the second and third modules resemble the natural activity
from the monkey's F5 and M1 regions, respectively. The network dynamics show a number of
other correspondences to the monkeys' natural brain activity as well, as detailed in the paper.

Notably, the simulated circuit's activity shows a relatively clear separation of the
object shapes.  That is - the visual information input to the network differentiates one trial
from another, and the circuit is able to leverage that information to properly condition
the hand shape trajectory for grasping the object of that trial's particular shape, size, and location.
The visual information forward-propagates through the network, through successive processing
steps, where it is eventually leveraged to recapitulate the grasp appropriate for the object
represented by the input visual features. As we will note below: in the absense of this
visual information forward propagating, the circuit can at-best learn a loss-minimizing grasp,
stereotyped across all object sizes and shapes. Such a grasp bears some resemblance to all
grasps the given monkey performed, due to all trials being a reach-to-grasp preceded by a
waiting period, but performance suffers drastically.

As a result, simulating a brain lesion in terms of this artificial network results in error modes
which resemble a natural lesion of certain regions of a primate brain.

...
* Lesioning properties

* Long-running dynamics

@@

@@

* Overview of test bed
** Michaels model .
** Lesion designs
*** Lesion examples, i.e. hand velocity more affected than shoulder velocity.
*** Disconnect modules vs lesion M1, and the effects
** Stimulation design
*** Spatio-temporal smoothing
** Observation model
